Urbanization and Public Transportation of Los Angeles

### 2023-11-09 Paper.Rmd
rm
Update Paper.Rmd

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(latex2exp)
library(dplyr)
library(tidyr)
library(scales)
library(stargazer)
ridership_data <- read.csv("data/April_Transit_Ridership.csv")
budget_data <- read.csv("data/City_Budget_Expenditures_on_Transportation_20231106.csv")
```

@@

@@

# Create the line plot
ggplot(ridership_long, aes(x=Year, y=RateOfChange, color=Type, group=Type)) +
  geom_line() +
  scale_y_continuous(labels = label_comma()) +
  labs(title="Trend in Rate of Change of Public Transportation Usage",
       x="Year",
       y="Rate of Change in Total Monthly Ridership",
       color="Transportation Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(2019, 2023))
```

@@

The following R code calculates the rate of change for Total Monthly Bus and Total Monthly Rail ridership and performs simple linear regression on these rates to understand how they change over the years:

```{r rate-of-change}
# Calculate the rate of change for Total Monthly Bus and Rail
ridership_data$BusRateOfChange <- c(NA, diff(ridership_data$Total.Monthly.Bus))
ridership_data$RailRateOfChange <- c(NA, diff(ridership_data$Total.Monthly.Rail))

@@

```{r rate-of-change-results, results='asis'}
stargazer(bus_model, rail_model, type = "latex", title = "Linear Regression Results", header=FALSE, omit=c("Year","Constant"))
stargazer(bus_model_poly, rail_model_poly, type = "latex", title = "Polynomial Regression Results", header=FALSE, omit=c("poly","Constant"))
```

```{r predict-trend-usage}
ridership_trend_long <- gather(ridership_data, key = "Type", value = "TotalRidership", -Year)

@@

@@

```{r trend-usage, fig.cap="Line plot of public transportation ridership over time with regression lines"}
# Create the line plot with observed data
p <- ggplot(ridership_trend_long, aes(x=Year, y=TotalRidership, color=Type, group=Type)) +
  geom_line() +
  # Assuming ridership_data has a 'Year' column and the predicted values only.
  # Add the regression lines for predicted bus data (explicit aesthetics)
  geom_line(data=ridership_data, aes(x=Year, y=PredictedBusROC, group=1), color="blue", linetype="dashed", size=1) +
  # Add the regression lines for predicted rail data (explicit aesthetics)
  geom_line(data=ridership_data, aes(x=Year, y=PredictedRailROC, group=1), color="red", linetype="dashed", size=1) +
  geom_line(data=ridership_data, aes(x=Year, y=PredictedBusROC_poly, group=2), color="hotpink", linetype="dotted", size=1) +
  # Add the regression lines for predicted rail data (explicit aesthetics)
  geom_line(data=ridership_data, aes(x=Year, y=PredictedRailROC_poly, group=2), color="darkgreen", linetype="dotted", size=1) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(title="Trend in Public Transportation Usage",
       x="Year",
       y="Total Monthly Ridership",
       color="Transportation Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(2019, 2023))

# Data of 2023
last_year <- max(ridership_data$Year)

# Add labels for the dashed lines
p + annotate("text", x=last_year, y=ridership_data$PredictedBusROC[ridership_data$Year == last_year], label="Predicted Bus Linear", color="blue", hjust=1.1, vjust=6) +
  annotate("text", x=last_year, y=ridership_data$PredictedRailROC[ridership_data$Year == last_year], label="Predicted Rail Linear", color="red", hjust=1.1, vjust=6) +
  annotate("text", x=last_year, y=ridership_data$PredictedRailROC_poly[ridership_data$Year == last_year], label="Predicted Bus Quadratic", color="hotpink", hjust=1.1, vjust=8) +
  annotate("text", x=last_year, y=ridership_data$PredictedRailROC_poly[ridership_data$Year == last_year], label="Predicted Rail Quadratic", color="darkgreen", hjust=1.1, vjust=10)
```

>>>>

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(latex2exp)
library(dplyr)
library(tidyr)
library(scales)
library(stargazer)
ridership_data <- read.csv("data/April_Transit_Ridership.csv")
population_data <- read.csv("data/Los_Angeles_Population.csv")
budget_data <- read.csv("data/City_Budget_Expenditures_on_Transportation_20231106.csv")
```

@@

Testhfgg

@@

# Create the line plot
ggplot(ridership_long, aes(x=Year, y=RateOfChange, color=Type, group=Type)) +
  geom_line() +
  scale_y_continuous(labels = label_comma()) +
  labs(title="Trend in Rate of Change of Public Transportation Usage",
       x="April of Each Year",
       y="Rate of Change in Total Monthly Ridership",
       color="Transportation Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(2019, 2023))
```

@@

```{r rate-of-change-and-regression}
# Derivative Analysis and Regression
# Calculate the rate of change for Total Monthly Bus and Rail
ridership_data$BusRateOfChange <- c(NA, diff(ridership_data$Total.Monthly.Bus))
ridership_data$RailRateOfChange <- c(NA, diff(ridership_data$Total.Monthly.Rail))

@@

@@

# Normalize the ridership data with the population data to get ridership per capita
ridership_data <- ridership_data %>%
  left_join(population_data, by = "Year") %>%
  mutate(
    RidershipPerCapitaBus = Total.Monthly.Bus / population,
    RidershipPerCapitaRail = Total.Monthly.Rail / population
  )

# Calculate the rate of change for Ridership per Capita
ridership_data$BusRateOfChangePerCapita <- c(NA, diff(ridership_data$RidershipPerCapitaBus))
ridership_data$RailRateOfChangePerCapita <- c(NA, diff(ridership_data$RidershipPerCapitaRail))

# Linear and polynomial regression for ridership per capita
bus_model_per_capita <- lm(BusRateOfChangePerCapita ~ Year, data = ridership_data)
rail_model_per_capita <- lm(RailRateOfChangePerCapita ~ Year, data = ridership_data)
```

```{r regression-results, results='asis'}
# Regression Results
stargazer(bus_model, rail_model, type = "latex", title = "Linear Regression Results", header=FALSE, omit=c("Year","Constant"))
stargazer(bus_model_poly, rail_model_poly, type = "latex", title = "Polynomial Regression Results", header=FALSE, omit=c("poly","Constant"))
```

```{r predict-trend-usage}
# Predict the rate of change per capita for all years using the linear and polynomial models
# (This assumes you have a model fitted to per capita data)
all_years <- data.frame(Year = seq(min(ridership_data$Year), 2027))
all_years <- left_join(all_years, population_data, by = "Year")

all_years$PredictedBusROCPerCapita <- predict(bus_model_per_capita, newdata=all_years)
all_years$PredictedRailROCPerCapita <- predict(rail_model_per_capita, newdata=all_years)

# Assuming the rate of change per capita is per year, integrate the rate of change to get the total ridership per capita
first_known_bus_ridership_per_capita <- head(ridership_data$RidershipPerCapitaBus, 1)
first_known_rail_ridership_per_capita <- head(ridership_data$RidershipPerCapitaRail, 1)

all_years$RidershipPerCapitaBus <- first_known_bus_ridership_per_capita + cumsum(c(0, all_years$PredictedBusROCPerCapita[-1]))
all_years$RidershipPerCapitaRail <- first_known_rail_ridership_per_capita + cumsum(c(0, all_years$PredictedRailROCPerCapita[-1]))
```

```{r format-and-combine-data}
# Format and Combine Data
# Convert the predicted data to long format
predicted_data_long <- all_years %>%
  select(Year, RidershipPerCapitaBus, RidershipPerCapitaRail) %>%
  pivot_longer(cols = c("RidershipPerCapitaBus", "RidershipPerCapitaRail"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Predicted")

# Now, make sure the actual data is in the same long format with the same column names
actual_data_long <- ridership_data %>%
  select(Year, RidershipPerCapitaBus, RidershipPerCapitaRail) %>%
  pivot_longer(cols = c("RidershipPerCapitaBus", "RidershipPerCapitaRail"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Actual")

# Combine the actual and predicted data
combined_data_long <- rbind(actual_data_long, predicted_data_long)

# Filter out the extreme predicted values for the polynomial bus data
combined_data_long <- combined_data_long %>%
  filter(!(Type == "RidershipPerCapitaBus" & TotalRidership < -10000000 & Prediction == "Predicted"))

@@

```{r plot-trend-usage, fig.cap="Line plot of public transportation ridership over time with regression lines"}
# Create the line plot with observed and predicted per capita data
p <- ggplot(combined_data_long, aes(x = Year, y = RidershipPerCapita, color = Type)) +
  geom_line(aes(linetype = Prediction)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(title = "Trend in Public Transportation Usage Per Capita",
       x = "April of Each Year",
       y = "Ridership Per Capita",
       color = "Transportation Type",
       linetype = "Data Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(min(ridership_data$Year), max(all_years$Year)))

# Print the plot
print(p)

# Sources:

- Transit ridership data: Los Angeles Metro "L.A. METRO TRANSIT RIDERSHIP UP 10 PERCENT, SETS POST-PANDEMIC RECORD", Patrick Chandler

- Los Angeles Population Data (real and predicted): WorldPopulationReview "Los Angeles, California Population"
